\documentclass{beamer}
\usepackage{xcolor}
\usetheme{warsaw}
\useoutertheme{infolines}

\begin{document}
	
	\begin{frame}
	\textcolor{blue}{\textbf{Example 7.15}}\hspace{5mm}Construct a PDA to accept the following language L=$\lbrace$$a^{2n}b^{n}$,where $n>0$$\rbrace$.\\
	\textbf{Solution}: The language consists of 2n number of ‘a’ and n number of ‘b’ belong to the language set. The
	PDA can be designed as follows—at the time of traversing the fi rst ‘a’, two $z_1$ are added to the stack with
	a state change from $q_0$ to $q_1$ . At the time of traversing the second ‘a’, one $z_1$ is popped from the stack
	with a state change from $q_1$ to $q_0$ . By this process after traversing 2n number of ‘a’, only n number of $z_1$
	exist in the stack. At the time of traversing the first ‘b’, the stack top is $z_1$ and the state is $q_0$. Those $z_1$ are
	popped at the time of traversing n number of ‘b’.
	The transitional functions are
	\end{frame}
	\begin{frame}
		$\delta$$(q_0,a,z_0)\xrightarrow{}(q_1,z_1z_1z_0)$\\
		$\delta$$(q_1,a,z_1)\xrightarrow{}(q_0,\lambda)$\\
		$\delta$$(q_0,a,z_1)\xrightarrow{}(q_1,z_1z_1z_0)$\\
		$\delta$$(q_0,b,z_1)\xrightarrow{}(q_2,\lambda)$\\
		$\delta$$(q_2,b,z_1)\xrightarrow{}(q_2,\lambda)$\\
		$\delta$$(q_2,\lambda,z_0)\xrightarrow{}(q_f,z_0)$// Accepted by final state\\
		$\delta$$(q_2,\lambda,z_0)\xrightarrow{}(q_2,\lambda)$// Accepted by empty stack
	\end{frame}
\begin{frame}
	\textcolor{blue}{\textbf{Example 7.16}}\hspace{5mm}Construct a PDA to accept the language L =$\lbrace$$a^{n}b^{n}c^{m}$,where n, $m \geq 1$$\rbrace$ by the empty stack and by the fi nal state.\\
	\textbf{Solution}: The language is in the form $a^{n}b^{n}c^{m}$, where n, $m \geq 1$. In the language set, the number of ‘a’ and
	the number of ‘b’ are the same, but the number of ‘c’ is different. All the strings in the language set start
	with n number of ‘a’s followed by n number of ‘b’s and ends with m number of ‘c’s. Here, m and n are
	both $\geq$ 1, and the null string does not belong to the language set.
	The PDA for the language can be designed in the following way.
	When traversing the ‘a’s from the input tape, the $z_1$’s are pushed in the stack one by one. At the time
	of traversing the ‘b’s, all the $z_1$’s which were pushed into the stack are popped one by one. When the
	first ‘c’ will be traversed, at that time the machine is in state $q_1$ and stack top $z_0$. At the time of traversal
	of m number of ‘c’s, no operation is done on the stack.
	The transition function for constructing the PDA for the string $a^{n}b^{n}c^{m}$ where m, $n \geq 1$ are
\end{frame}
\begin{frame}
	$\delta$$(q_0,a,z_0)\xrightarrow{}(q_0,z_1z_0)$\\
	$\delta$$(q_0,a,z_1)\xrightarrow{}(q_0,z_1z_1)$\\
	$\delta$$(q_0,b,z_1)\xrightarrow{}(q_1,\lambda)$\\
	$\delta$$(q_1,b,z_1)\xrightarrow{}(q_1,\lambda)$\\
	$\delta$$(q_1,c,z_0)\xrightarrow{}(q_1,z_0)$\\
	$\delta$$(q_1,\lambda,z_0)\xrightarrow{}(q_1,\lambda)$ accepted by the empty stack\\
	$\delta$$(q_1,\lambda,z_0)\xrightarrow{}(q_f,z_0)$ accepted by the fi nal state\\
\end{frame}

	
\end{document}